---
title: "DS 2021 Jelly Figures"
author: "Naomi Lubkin and Mara Bohm"
date: "15/12/2021"
output: pdf_document
---

```{r, results=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=70), tidy=TRUE)
# code above ensures no text is cut off when knit
```

# Loading In Packages and Data

```{r, results=FALSE, message=FALSE, warning=FALSE}
library(patchwork)
library(car)
library(performance)
library(DHARMa)
library(fitdistrplus)
library(gamlss)
library(tidyverse)

polyp_data <- ds2021_polyp_data

```


# Cleaning the Data

```{r, results=FALSE, message=FALSE, warning=FALSE}

polyp_clean <- polyp_data %>%
  rename(collection_day = 'Data collection day', 
         jar_code = 'Jar Code (ex. E1)', 
         treatment = Chemical, 
         num_elongated = 'Elongation?', 
         num_ruffled = 'Ruffled?', 
         num_asexual_buds = 'Asexual Repro?', 
         total_num = Total,
         time_in_days = days_after) %>%
  mutate(jar_code = as.factor(jar_code),
         treatment = as.factor(treatment), 
         treatment = fct_relevel(treatment, 
                                 "Control", 
                                 "Caffeine", 
                                 "Estradiol", 
                                 "Combo"))  %>%
  dplyr::select(collection_day, 
                jar_code, 
                treatment, 
                num_elongated, 
                num_ruffled, 
                num_asexual_buds, 
                total_num,
                time_in_days) 

```

# Making Single Treatment Data Frames for Later Analyses

```{r, results=FALSE, message=FALSE, warning=FALSE}

# data frame with only control
only_control <- polyp_clean %>%
  filter(treatment == "Control")

# data frame with only caffeine
only_caf <- polyp_clean %>%
  filter(treatment == "Caffeine")

# data frame with only estradiol
only_est <- polyp_clean %>%
  filter(treatment == "Estradiol")

# data frame with only combination
only_combo <- polyp_clean %>%
  filter(treatment == "Combination")

# data frame with caffeine and control
just_caf <- polyp_clean %>%
  filter(treatment == "Caffeine" | treatment == "Control")

# data frame with combination and control
just_combo <- polyp_clean %>%
  filter(treatment == "Combination" | treatment == "Control")

# data frame with caffeine, combination, and control
just_caf_combo <- polyp_clean %>%
  filter(treatment == "Combination" | treatment == "Control" | treatment == "Caffeine")

```

# Figure 1: Response Variables Over Time

### Figure 1A: Ruffled Polyps Over Time
```{r, message=FALSE, warning=FALSE, fig.show='hide'}

ruffles_time <- ggplot(polyp_clean, aes(x=time_in_days,  y=num_ruffled)) +
  geom_point(colour="#fb9b06") +
  geom_smooth(method = "lm", se=FALSE, colour="black", show.legend=FALSE) +
  labs(x = "Time (days)", y = "Number of Ruffled Polyps") +
  scale_x_continuous(breaks=c(0, 3, 6, 9, 12)) +
  theme_classic() +
    theme(legend.position = "none") +
  ggtitle("A. Ruffling") 
  
# Spearman's rank correlation coefficient for the correlation between the number  of ruffled polyps and time
cor(polyp_clean$time_in_days, polyp_clean$num_ruffled, method = "spearman")

```

### Figure 1B: Elongated Polyps Over Time
```{r, message=FALSE, warning=FALSE, fig.show='hide'}

elo_time <- ggplot(polyp_clean, aes(x=time_in_days,  y=num_elongated)) +
  geom_point(colour="#e55c30") +
  geom_smooth(method = "lm", se=FALSE, colour="black", show.legend=FALSE) +
  labs(x = "Time (days)", 
       y = "Number of Elongated Polyps") +
  scale_x_continuous(breaks=c(0, 3, 6, 9, 12)) +
  theme_classic() +
    theme(legend.position = "none") +
  ggtitle("B. Elongation") 
  
# Spearman's rank correlation coefficient for the correlation between the number  of elongated polyps and time
cor(polyp_clean$time_in_days,polyp_clean$num_elongated, method= "spearman")

```

### Figure 1C: Asexual Buds Over Time
```{r, message=FALSE, warning=FALSE, fig.show='hide'}

buds_time <- ggplot(polyp_clean, aes(x=time_in_days,  y=num_asexual_buds)) +
  geom_point(colour="#450a69") +
  geom_smooth(method = "lm", se=FALSE, colour="black", show.legend=FALSE) +
  labs(x = "Time (days)", y = "Number of Asexual Buds") +
  scale_x_continuous(breaks=c(0, 3, 6, 9, 12)) +
  theme_classic() +
    theme(legend.position = "none") +
  ggtitle("C. Budding") 
  
# Spearman's rank correlation coefficient for the correlation between number of   asexual buds and time 
cor(polyp_clean$time_in_days,polyp_clean$num_asexual_buds, method= "spearman")

```

### Figure 1D: Total Number of Polyps Over Time
```{r, message=FALSE, warning=FALSE, fig.show='hide'}

total_time <- ggplot(polyp_clean, aes(x=time_in_days,  y=total_num)) +
  geom_point(colour="#b1325a") +
  geom_smooth(method = "lm", se=FALSE, colour="black", show.legend=FALSE) +
  labs(x = "Time (days)", y = "Total Number of Polyps") +
  scale_x_continuous(breaks=c(0, 3, 6, 9, 12))+
  theme_classic() +
    theme(legend.position = "none") +
  ggtitle("D. Total Number") 
  
# Spearman's rank correlation coefficient for the correlation between total number   of polyps and time 
cor(polyp_clean$time_in_days,polyp_clean$total_num, method= "spearman")

```

### Merging The Four Graphs Into One Figure Through Patchwork
```{r, results=FALSE, message=FALSE, warning=FALSE}

(ruffles_time | elo_time)/(buds_time | total_time)

```

# Figure 2: Ruffling In Each Treatment

### Building a Box Plot For Control, Caffeine, Estradiol and Combination with Time as a Factor
```{r, message=FALSE, warning=FALSE}

box_caf_combo<- ggplot(just_caf_combo, aes(  x = as.factor(time_in_days), y = num_ruffled, fill=treatment))+
  geom_boxplot() +
  theme_classic() +
  scale_fill_manual(values = c("#404788FF", "#238A8DFF","#55C667FF")) +
  labs(x = "Time (days)", y = "Number of Ruffled Polyps", fill ="Treatment") 
box_caf_combo


#Spearman's rank correlation coefficient for the correlation between number of  ruffling polyps and time in each treatment 

# control: 
cor(only_control$time_in_days, only_control$num_elongated, method= "spearman")

#caffeine 
cor(only_control$time_in_days, only_caf$num_elongated, method= "spearman")

#estradiol 
cor(only_control$time_in_days, only_est$num_elongated, method= "spearman")

# combination: 
cor(only_combo$time_in_days, only_combo$num_elongated, method= "spearman")



```


# Figure 3: Elongation per treatment and over time

# Figure 3A: Box Plot of Overall Elongation per Treatment
```{r, message=FALSE, warning=FALSE, fig.show='hide'}

box_elongation <- ggplot(polyp_clean, aes(x=treatment, y=num_elongated, fill = treatment)) +
  geom_boxplot() +
  theme_classic() +
  scale_fill_manual(values = c("#404788FF", "#238A8DFF", "#FDE725FF","#55C667FF")) +
  labs(x = "Treatment", y = "Number of Elongated Polyps") +
  theme(legend.position = "none", axis.text.x = element_text(angle=30, vjust = 0.8, hjust = 0.7)) +
  ggtitle("A")
box_elongation

```

### Figure 3B: Scatter/Line Plot of the Number of Elongated Polpys over Time, Per Treatment
```{r, message=FALSE, warning=FALSE, fig.show='hide'}

plot_elongation <- ggplot(polyp_clean, 
                          aes(x=time_in_days, 
                              y=num_elongated, 
                              colour=treatment, 
                              shape = treatment))+
  geom_point(aes(x=time_in_days, 
                 y=num_elongated, 
                 colour=treatment, 
                 shape = treatment)) +
  theme_classic() +
  geom_smooth(se=FALSE, method="lm") +
  labs(x = "Time (days)", 
       y = "Number of Elongated Polyps", 
       colour ="Treatment", 
       shape = NULL) +
  scale_colour_manual(name = "Treatment", 
                      labels= c("Control", 
                                "Caffeine", 
                                "Estradiol", 
                                "Combination"), 
                      values = c("#404788FF", 
                                 "#238A8DFF",  
                                 "#FDE725FF",
                                 "#55C667FF")) +
  scale_shape_manual(name = "Treatment", 
                     labels= c("Control", 
                               "Caffeine", 
                               "Estradiol", 
                               "Combination"), 
                     values = c(15,19,17,8)) +
  scale_x_continuous(breaks=c(0, 3, 6, 9, 12)) +
  ggtitle("B")


# Spearman's Rank Correlation Coefficient for Correlation Between Number of   Elongated Polyps and Time in Each Treatment

# control: 
cor(only_control$time_in_days,only_control$num_elongated, method= "spearman")

# caffeine: 
cor(only_caf$time_in_days,only_caf$num_elongated, method= "spearman")

# estradiol: 
cor(only_est$time_in_days,only_est$num_elongated, method= "spearman")

# combination: 
cor(only_combo$time_in_days,only_combo$num_elongated, method= "spearman")

```
  
### Merging The Two Graphs Into One Figure Through Patchwork

```{r, message=FALSE, warning=FALSE}

(box_elongation | plot_elongation)

```

# Figure 4: Budding in the Control and Combination Treatments

### Figure 4A: Box Plot of Overall Budding in Control and Combo
```{r, message=FALSE, warning=FALSE, fig.show='hide'}

box_buds <-ggplot(just_combo, aes(x=treatment, y=num_asexual_buds, fill = treatment)) +
  geom_boxplot() +
  theme_classic() +
  scale_fill_manual(values = c("#404788FF", "#55C667FF")) +
  labs(x = "Treatment", y = "Number of Asexual Buds") +
  theme(legend.position = "none")  +
   ggtitle("A")

```

### Figure 4B: Scatter/Line Plot of Budding in Control and Combination over Time
```{r, message=FALSE, warning=FALSE, fig.show='hide'}

plot_buds <- ggplot(just_combo, aes(x=time_in_days, y=num_asexual_buds, colour=treatment, shape = treatment)) +
  geom_point() +
  theme_classic() +
   labs(x = "Time (days)", y = "Number of Asexual Buds", colour ="Treatment") +
  geom_smooth(se= FALSE, method="lm")+
   scale_colour_manual(name = "Treatment", labels= c("Control", "Combination"), values = c("#404788FF", "#55C667FF")) +
  scale_shape_manual(name = "Treatment", labels= c("Control", "Combination"), values = c(15,8)) +
  scale_x_continuous(breaks=c(0, 3, 6, 9, 12)) +
  ggtitle("B")
  
#Spearman's rank correlation coefficient for the correlation between number of  asexual buds and time in each treatment 

# control: 
cor(only_control$time_in_days, only_control$num_asexual_buds, method= "spearman")

#caffeine 
cor(only_control$time_in_days, only_caf$num_asexual_buds, method= "spearman")

#estradiol 
cor(only_control$time_in_days, only_est$num_asexual_buds, method= "spearman")

# combination: 
cor(only_combo$time_in_days, only_combo$num_asexual_buds, method= "spearman")



```

### Merging The Two Graphs Into One Figure Through Patchwork

```{r, message=FALSE, warning=FALSE}

(box_buds | plot_buds)

```

# Figure 4: Total Number of Polyps

### Figure 4A: Box Plot for Overall Total Number of Polyps by Treatment
```{r, message=FALSE, warning=FALSE, fig.show='hide'}

box_total <- ggplot(polyp_clean, aes(x=treatment, y=total_num, fill = treatment)) +
  geom_boxplot() +
  theme_classic() +
  scale_fill_manual(values = c("#404788FF", "#238A8DFF", "#FDE725FF","#55C667FF")) +
  labs(x = "Treatment", y = "Total Number of  Polyps") +
  theme(legend.position = "none", axis.text.x = element_text(angle=30, vjust = 0.8, hjust = 0.7)) + 
  ggtitle("A")

```  

### Figure 4B: Scatter/Line Plot of Total Number of Polyps Over Time, Broken Up By Treatment
```{r, message=FALSE, warning=FALSE, fig.show='hide'}

plot_total <- ggplot(polyp_clean, aes(x=time_in_days, y=total_num, colour=treatment, shape = treatment))+
  geom_point(aes(colour=treatment)) +
  theme_classic() +
  geom_smooth(se=FALSE, method="lm") +
  labs(x = "Time (days)", y = "Total Number of Polyps", colour ="Treatment") +
  scale_colour_manual(name = "Treatment", labels= c("Control", "Caffeine", "Estradiol", "Combination"), values = c("#404788FF", "#238A8DFF",  "#FDE725FF","#55C667FF")) +
  scale_shape_manual(name = "Treatment", labels= c("Control", "Caffeine", "Estradiol", "Combination"), values = c(15,19,17,8)) +
  scale_x_continuous(breaks=c(0, 3, 6, 9, 12)) +
  ggtitle("B")

# Spearman's rank correlation coefficient for the correlation between total   number of polyps and time in each treatment 
# control: 
cor(only_control$time_in_days,only_control$total_num, method= "spearman")

# caffeine: 
cor(only_caf$time_in_days,only_caf$total_num, method= "spearman")

# estradiol: 
cor(only_est$time_in_days,only_est$total_num, method= "spearman")

# combination: 
cor(only_combo$time_in_days,only_combo$total_num, method= "spearman")

```

### Merging The Two Graphs Into One Figure Through Patchwork
```{r, message=FALSE, warning=FALSE}

(box_total | plot_total)
```

